Type theories as quotient inductive-recursive types

### 2025-02-21 tex/main.tex
Number the section on the elimination rule in appendix

We refer to \cref{appendix:full-def} and our \Agda development for the full definition of the elimination principle.

@@

\paragraph*{Type theory as a quotient inductive-inductive-recursive type}
We give the QIIRT for type theory with a universe of small types, $\Pi$-types, and extensional identity types in one place below.

@@

\begin{alignat*}{5}
  \pi_2\beta      & : \implicit{\Gamma, \Delta, \Theta, \sigma, A, t}\; && \pi_2(\sigma , t)        && =^{\Tm\;\Gamma\; A} &&  t \\
  [\idS]         & : \implicit{\Gamma, A, t}\; && {\sub{t}{\idS}}          && =^{\Tm\;\Gamma\;A} && t \\
  [;]            & :\implicit{\Gamma, \Delta, \Theta, \sigma, \tau, t}\; && {\sub{t}{\sigma ; \tau}} && =^{\Tm\;\Gamma\;\subTy{A}{\sigma ; \tau}} && {\sub{(\sub{t}{\tau})}{\sigma}} \\
  []\lambda       & : \implicit{\Gamma, \Delta, \sigma, A, t}\;&& \sub{(\lambda\,t)}{\sigma} && {} =^{\Tm\,\Gamma\,(\subTy{(\Pi\,A\,B)}{\sigma})} {} && \lambda (\sub{t}{\sigma \uparrow A}) \\
  \Pi\beta        & : \implicit{\Gamma, A, B, t}\; && \mathsf{app}\,(\lambda\,t)            && =^{\Tm\,(\Gamma, A)\,B} && t \\
  \Pi\eta         & : \implicit{\Gamma, A, B, t}\; && t                                     && =^{\Tm\,\Gamma\,(\Pi\,A\,B)} && \lambda(\mathsf{app}\,t) \\
  []\mathsf{refl} & : \implicit{\Gamma, a, t}\;&& \sub{(\mathsf{refl}\;t)}{\sigma} && =^{\Tm\;\Gamma\;(\subTy{(\mathsf{Id}\,a\,t\,t)}{\sigma})} {} && \mathsf{refl}(\sub{t}{\sigma})
\end{alignat*}

@@

\paragraph*{Elimination principles for type theories as QIIRTs}

\subparagraph*{Motives}
\begin{align*}
  \Ctx^M & : \Ctx \to \Set &
  \Sub^M & : \implicit{\Gamma, \Delta}\,\Ctx^M\;\Gamma \to \Ctx^M\;\Delta \to \Sub\;\Gamma\;\Delta \to \Set \\
  \Ty^M  & : \implicit{\Gamma}\,\Ctx^M\;\Gamma \to \Ty\;\Gamma \to \Set &
  \Tm^M  & : \implicit{\Gamma, A}\; (\Gamma^M : \Ctx^M\;\Gamma) \to \Ty^M\;\Gamma^M\;A \to \Tm\;\Gamma\;A \to \Set
\end{align*}

\subparagraph*{Methods}
Note that implicitly quantified variables which would be otherwise declared inside $\implicit{\dots}$ are omitted below.
Each variable $\cc^M$ with the superscript $(\blank)^M$ is always indexed by $\cc$ in addition to other indices.

>>>>

We refer to \cref{subsec:def:elimination} for the full definition of the elimination principle.

@@

\subsection{Type theory as a quotient inductive-inductive-recursive type}
We give the QIIRT for type theory with a universe of small types, $\Pi$-types, and extensional identity types in one place below.

@@

\begin{alignat*}{5}
  \pi_2\beta      & : \implicit{\Gamma, \Delta, \Theta, \sigma, A, t}\;  & \pi_2(\sigma , t)                & =^{\Tm\;\Gamma\; A} &&  t \\
  [\idS]         & : \implicit{\Gamma, A, t}\;                           & {\sub{t}{\idS}}                  & =^{\Tm\;\Gamma\;A} && t \\
  [;]            & :\implicit{\Gamma, \Delta, \Theta, \sigma, \tau, t}\; & {\sub{t}{\sigma ; \tau}}         & =^{\Tm\;\Gamma\;\subTy{A}{\sigma ; \tau}} && {\sub{\sub{t}{\tau}}{\sigma}} \\
  []\lambda       & : \implicit{\Gamma, \Delta, \sigma, A, t}\;          & \sub{(\lambda\,t)}{\sigma}       & =^{\Tm\,\Gamma\,(\subTy{(\Pi\,A\,B)}{\sigma})} {} && \lambda (\sub{t}{\sigma \uparrow A}) \\
  []\mathsf{refl} & : \implicit{\Gamma, a, t}\;                          & \sub{(\mathsf{refl}\;t)}{\sigma} & =^{\Tm\;\Gamma\;(\subTy{(\mathsf{Id}\,a\,t\,t)}{\sigma})} {} && \mathsf{refl}\,(\sub{t}{\sigma}) \\
  \Pi\beta        & : \implicit{\Gamma, A, B, t}\;                       & \mathsf{app}\,(\lambda\,t)       & =^{\Tm\,(\Gamma, A)\,B} && t \\
  \Pi\eta         & : \implicit{\Gamma, A, B, t}\;                       & t                                & =^{\Tm\,\Gamma\,(\Pi\,A\,B)} && \lambda(\mathsf{app}\,t)
\end{alignat*}

@@

\subsection{The elimination principle for type theory as QIIRT} \label{subsec:def:elimination}

\paragraph*{Motives}
\begin{align*}
  \Ctx^M & : \Ctx \to \Set &
  \Sub^M & : \implicit{\Gamma, \Delta}\,\Ctx^M\;\Gamma \to \Ctx^M\;\Delta \to \Sub\;\Gamma\;\Delta \to \Set \\
  \Ty^M  & : \implicit{\Gamma}\,\Ctx^M\;\Gamma \to \Ty\;\Gamma \to \Set &
  \Tm^M  & : \implicit{\Gamma, A}\; (\Gamma^M : \Ctx^M\;\Gamma) \to \Ty^M\;\Gamma^M\;A \to \Tm\;\Gamma\;A \to \Set
\end{align*}

\paragraph*{Methods}
Note that implicitly quantified variables which would be otherwise declared inside $\implicit{\dots}$ are omitted below.
Each variable $\cc^M$ with the superscript $(\blank)^M$ is always indexed by $\cc$ in addition to other indices.
